---
title: "ESTP-course in BI with Power BI"
description: |
  Information about the ESTP-webinar on Power BI held on the 22.-22. april 2021.  
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE}
library(metathis)
library(magrittr)

meta() %>%
  meta_description(
    "ESTP-course on Power BI in april of 2021."
  ) %>% 
  meta_name("github-repo" = "skars82/estp_course_pbi") %>% 
  meta_viewport() %>% 
  meta_social(
    title = "Introduction to business intelligence in public statistics (Power BI)",
    url = "https://skars82.github.io/estp_course_pbi/",
    image = "https://github.com/skars82/slides/blob/master/docs/estp_course_powerbi/pictures/twitter_slide.png)",
    image_alt = "Title slide of first day",
    og_type = "Course",
    og_author = c("Øyvind Bruer-Skarsbø", "Carl Franklin Corneil"),
    twitter_card_type = "summary",
    twitter_creator = "@YSkarsb"
  )

```


# Introduction

This course/webinar is intended for newcomers to Power BI, who want to learn more about how to use
a business intelligence tool like Power BI to visualize data. The course is also geared towards the 
type of data ususally seen in the production of official statistics.  


# 1. Installation

The only prerequisite for attending the course is that you have [Power BI Desktop installed on a computer](https://powerbi.microsoft.com/en-us/downloads/) that 
is at your disposal during the 3-day course. We will only use the Desktop-software, which is free, so there
is no need to install anything more. PLease download the latest version.

## 1.2 Settings

We recommend to choose the following settings in Power BI desktop: 

- Set english as the language. 
- Security > Allow Data Extensions

# 2. Getting started

## 2.1 Introduction video

This video gives you a nice introduction to what Power BI is:

<iframe width="560" height="315" src="https://www.youtube.com/embed/Vqz2d7pTOV8" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

# 3. Day 1

## 3.1 Fundamentals Power BI and data visualization
- Power BIs platform for sharing reports
- Pareto principle - COCOMO (wikipedia)
- Download powerpoint: 

## 3.2 Importing data and creating a report
<a href="../data/companies_2020_10000_v001.csv">CSV-download</a>

<iframe src="https://player.vimeo.com/video/305761177?color=90cc93&portrait=0" width="640" height="360" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

## 3.3 Build data model with more data
Nace-code descriptions from Eurostat: <a href="https://ec.europa.eu/eurostat/web/nace-rev2/correspondence_tables"> Excelfile </a>


<iframe src="https://player.vimeo.com/video/305761177?color=90cc93&portrait=0" width="640" height="360" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

# 5. Day 3 

## 5.1 Import data from Eurostat API using a custom connector

<iframe src="https://player.vimeo.com/video/305761177?color=90cc93&portrait=0" width="640" height="360" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

- Download the connector .mez-file: <a href="https://github.com/eurostat/PowerBI-connector/blob/master/Eurostat%20API.mez">press download on this page"</a>
- Make a folder here: 'C:\\Users\\[username]\\Documents\\Microsoft Power BI Desktop\\Custom Connectors'
- Add the .mez-file in the folder
- Change setting in Power BI > Security > Allow Data Extensions, restart Power BI




## 5.2 Import data from SSBs APIs

### 5.2.1 Import data from KLASS

<iframe src="https://player.vimeo.com/video/305761177?color=90cc93&portrait=0" width="640" height="360" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

<a href="https://www.ssb.no/en/klass/">ssb.no/en/klass/</a>

- Browse to the variant or correspondance you are after
- Note id at the end of the URL
- Change the code to reflect the ID, or maybe klassType to "variants"
- Check the documentation of the API: <a href="https://data.ssb.no/api/klass/v1/api-guide.html">data.ssb.no/api/klass"</a>

#### "Get data > Blank query > Advanced Editor":
<script src="https://gist.github.com/aecorn/a3f42cc730f12d6b8070ebf6e7d1da1e.js"></script>
### 5.2.2 Import data from STATBANK

<iframe src="https://player.vimeo.com/video/305761177?color=90cc93&portrait=0" width="640" height="360" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

<a href="https://www.ssb.no/en/statbank">ssb.no/en/statbank</a>

- Find the ID of the table you want to get info from
- Expand the "API-query box" at the bottom of the page after following the whole web-interface.
- You might have to replace single double-quotes, with double double-quotes.

#### "Get data > Blank query > Advanced Editor":
<script src="https://gist.github.com/aecorn/0adf2cf352a889a769bb5f63c6a59a0d.js"></script>



