---
title: "ESTP-course in BI with Power BI"
description: |
  Information about the ESTP-webinar on Power BI held on the 20.-22. april 2021.  
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE}
library(metathis)
library(magrittr)

meta() %>%
  meta_description(
    "ESTP-course on Power BI in april of 2021."
  ) %>% 
  meta_name("github-repo" = "skars82/estp_course_pbi") %>% 
  meta_viewport() %>% 
  meta_social(
    title = "Introduction to business intelligence in public statistics (Power BI)",
    url = "https://skars82.github.io/estp_course_pbi/",
    image = "https://github.com/skars82/slides/blob/master/docs/estp_course_powerbi/pictures/twitter_slide.png)",
    image_alt = "Title slide of first day",
    og_type = "Course",
    og_author = c("Øyvind Bruer-Skarsbø", "Carl Franklin Corneil"),
    twitter_card_type = "summary",
    twitter_creator = "@YSkarsb"
  )

```


## Introduction

This course/webinar is intended for newcomers to Power BI, who want to learn more about how to use
a business intelligence tool like Power BI to visualize data. The course is also geared towards the 
type of data ususally seen in the production of official statistics.  


## 1. Installation

The only prerequisite for attending the course is that you have [Power BI Desktop installed on a computer](https://powerbi.microsoft.com/en-us/downloads/) that 
is at your disposal during the 3-day course. We will only use the Desktop-software, which is free, so there
is no need to install anything more. Please download the latest version.

### 1.2 Recommended settings

We recommend to choose the following settings in Power BI desktop: 

- Set english as the language. 
- Security > Allow Data Extensions

## 2. Resources

### 2.1 What is Power BI?

This video gives you a nice introduction to what Power BI is:

<iframe width="560" height="315" src="https://www.youtube.com/embed/Vqz2d7pTOV8" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

### 2.2 Importing data

Importing data is fundamental step in data-tool. Power BI makes this very easy and can connect a large amount of data sources: 

**Importing csv-files**
<iframe width="560" height="315" src="https://www.youtube.com/embed/wuLnv3QJCHg" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

**Example with a SQL server database (with parameters)**
<iframe width="560" height="315" src="https://www.youtube.com/embed/1UkDaiRGShM" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


### 2.3 Building a data model

When you start use Power BI you should try to understand the recommended way to organize you data into a model. Its a fundamental part of the Power BI workflow, and avoiding could case bad performance and/or errors in your results. The recommended way to build up a data model in Power BI is called *star schema*: 

**Whats the deal with star schemas in Power BI?**
<iframe width="560" height="315" src="https://www.youtube.com/embed/vZndrBBPiQc" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

A common problem for newcomers to Power BI is understand how to handle *dates*! Basically, the answer amounts to including a separat date-dimension-table. This is such a fundamental part of any Power BI workflow that I would recommend setting up a template project that prompts for the start- and -stopdates of your datetable. Then you never have to think about it again!! While your at it, you can also disable the automatic time intelligence in the settings. 

**How to create a datetable?**
<iframe width="560" height="315" src="https://www.youtube.com/embed/Pzn2B4laK9A" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Since I mentioned template-projects and parameters, Im also gonna share a video that shows how to set ut up: 

**Parameters and template-projects**
<iframe width="560" height="315" src="https://www.youtube.com/embed/JlUfz18cx_w" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


### 2.4 Themes and sharing JSON-files

Colors, fonts and other design-elements are inevitably a big part visualizing data, and are usually subject to stricts rules from your organization. We recommend that apply the desired design-principles using json-files in Power BI. Its really easy and very fast. Check out this video that Carl made:

**Themes with json**
<iframe src="https://player.vimeo.com/video/537137641?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>
<a href="https://docs.microsoft.com/en-us/power-bi/create-reports/desktop-report-themes">Official documentation: Use report themes in Power BI Desktop</a>

### 2.5 Import data from Eurostat API using a custom connector

We just found out about this trick a few weeks ago, through [Eurostats Github-page](https://github.com/eurostat/PowerBI-connector). Please take advantage of all the great data thats in these databases. Again, Carl shows you how:

<iframe src="https://player.vimeo.com/video/536746730?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>


- Download the connector .mez-file: <a href="https://github.com/eurostat/PowerBI-connector/blob/master/Eurostat%20API.mez" target="_blank" >press download on this page"</a>
- Make a folder here: 'C:\\Users\\[username]\\Documents\\Microsoft Power BI Desktop\\Custom Connectors'
- Add the .mez-file in the folder
- Change setting in Power BI > Security > Allow Data Extensions, restart Power BI

<!--
### 2.7 Import NACE-codelist from  Eurostat
-->


### 2.7 Import data from SSBs APIs - Import data from KLASS

<iframe src="https://player.vimeo.com/video/537588889?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

<a href="https://www.ssb.no/en/klass/">ssb.no/en/klass/</a>

- Browse to the variant or correspondance you are after
- Note id at the end of the URL
- Change the code to reflect the ID, or maybe klassType to "variants" / "versions" etc.
- Check the documentation of the API: <a href="https://data.ssb.no/api/klass/v1/api-guide.html">data.ssb.no/api/klass"</a>


<details>
<summary>Code to paste under: "Get data > Blank query > Advanced Editor"</summary>
<script src="https://gist.github.com/aecorn/a3f42cc730f12d6b8070ebf6e7d1da1e.js"></script>
</details>

### 2.8 Import data from STATBANK

<iframe src="https://player.vimeo.com/video/536752553?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

<iframe src="https://player.vimeo.com/video/536754342?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

<a href="https://www.ssb.no/en/statbank">ssb.no/en/statbank</a>

- Find the ID of the table you want to get info from
- Expand the "API-query box" at the bottom of the page after following the whole web-interface.
- You might have to replace single double-quotes, with double double-quotes.

<details>
<summary>Code to paste under: Get data > Blank query > Advanced Editor</summary>
<script src="https://gist.github.com/aecorn/0adf2cf352a889a769bb5f63c6a59a0d.js"></script>
</details>

<iframe src="https://player.vimeo.com/video/536743890?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

<details>
<summary>Working example of Python-script for getting a statbank table.</summary>
<script src="https://gist.github.com/aecorn/e833246f39255df94e3876c467b3046b.js"></script>
</details>



### 2.9 Visualize with Python and Seaborn
<iframe src="https://player.vimeo.com/video/536747489?color=90cc93&portrait=0" width="704" height="396" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>

